### 2021-01-01 highlights.tex
Updates from Overleaf

\documentclass[preprint,12pt]{elsarticle}

\begin{document}

Highlights
\begin{itemize}
    \item The R script program could automatically scan protein-coding genes and generate the Kaplan-Meier plots and Cox's univariate/multivariate tables.
    \item Using a serial cut from 30\% to 70\% percentile of the cohort, it could find the least P-value cutoff of the Kaplan-Meier analysis for each gene expression.
    \item Our analysis could discover the pronounced biomarkers, which impact HNSCC's survival under the stringent Bonferroni adjustment.
\end{itemize}

\end{document}